LACRE 2022: abstract

### 2022-01-13 docs/02_vocabulary.Rmd
Sample posterior

# vocabulary
participants %>% 
	drop_na(lp) %>%
	filter(valid_participant) %>% 
	ggplot() +
	facet_grid(location~age_group) +
	# geom_violin(
	# 	position = position_dodge(width = 0.75),
	# 	aes(lp, vocab_size_l1, colour = lp, fill = lp)
	# ) + 
	geom_hline(yintercept = 0.5, colour = "white") +
	geom_point(
		aes(lp, vocab_size_l1),
		shape = 1, colour = "white", size = 1, stroke = 1,
		position = position_jitter(width = 0.1, seed = 888)
	) +
	geom_boxplot(
		aes(lp, vocab_size_l1, colour = lp, fill = lp),
		width = 0.25, outlier.colour = NA, fill = "#0D1117",
		position = position_dodge(width = 0.75), size = 1
	) +
	geom_text(
		data = sample_size,
		aes(x = 1.5, y = 0.95, label = paste0("n = ", n)),
		colour = "white", fill = "white", show.legend = FALSE
	) +
	labs(y = "Receptive vocabulary (L1)", colour = "Group", fill = "Group") +
	scale_y_continuous(limits = c(0, 1), labels = percent) +
	theme(
		legend.title = element_blank(),
		legend.position = "none",
		panel.grid.major.y = element_blank(),
		axis.title.x = element_blank(),
		axis.text = element_text(size = 9)
	)

>>>>

# vocabulary
participants %>% 
	drop_na(lp) %>%
	filter(valid_participant) %>% 
	ggplot() +
	facet_grid(location~age_group) +
	# geom_violin(
	# 	position = position_dodge(width = 0.75),
	# 	aes(lp, vocab_size_l1, colour = lp, fill = lp)
	# ) + 
	geom_hline(yintercept = 0.5, colour = "white") +
	geom_point(
		aes(lp, vocab_size_l1),
		shape = 1, colour = "white", size = 1, stroke = 1,
		position = position_jitter(width = 0.1, seed = 888)
	) +
	geom_boxplot(
		aes(lp, vocab_size_l1, fill = lp),
		width = 0.25, outlier.colour = NA,
		position = position_dodge(width = 0.75), size = 1
	) +
	geom_text(
		data = sample_size,
		aes(x = 1.5, y = 0.95, label = paste0("n = ", n)),
		colour = "white", fill = "white", show.legend = FALSE
	) +
	labs(y = "Receptive vocabulary (L1)", colour = "Group", fill = "Group") +
	scale_y_continuous(limits = c(0, 1), labels = percent) +
	theme(
		legend.title = element_blank(),
		legend.position = "none",
		panel.grid.major.y = element_blank(),
		axis.title.x = element_blank(),
		axis.text = element_text(size = 9)
	)

### 2022-01-13 docs/06_results-bcn.Rmd
Sample posterior

# load objects
tar_load_globals()
tar_load(participants)
tar_load(attrition)
tar_load(vocabulary)
tar_load(gaze)
tar_load(model_fits)

@@

participants <- participants %>% 
	left_join(distinct(attrition, participant, age_group, valid_participant)) %>% 
	drop_na(participant) %>% 
	left_join(vocabulary) %>% 
	mutate(vocab_size_l1_center = scale(vocab_size_l1)[,1])

@@

```{r fixed_effects_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
draws <- gather_draws(model_fits$fit_bcn, `b_.*`, `sigma`, regex = TRUE)

@@

draws %>% 
	filter(!str_detect(.variable, "time_bin|sigma")) %>% 
	mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4)) %>% 
	ggplot(aes(.value, .variable, fill = stat(cut_cdf_qi(cdf, labels = percent, .width = c(0.50, 0.80, 0.95))))) +
	geom_vline(xintercept = 0, colour = "black") +
	stat_halfeye(point_color = "black", interval_color = "black") +
	labs(x = "Value", y = "Variable", fill = "CrI") +
	scale_fill_brewer(direction = -1, na.translate = FALSE) +
	theme(
		axis.title = element_blank(),
		legend.position = "top"
	)

@@

vcov(model_fits$fit_bcn) %>% 
	cov2cor() %>% 
	as.data.frame() %>% 
	rownames_to_column("term1") %>% 
	pivot_longer(-term1, names_to = "term2", values_to = ".value") %>% 
	filter(!grepl("time_bin", term1), !grepl("time_bin", term2)) %>% 
	ggplot(aes(fct_rev(term1), term2, fill = .value)) +
	geom_tile() +
	labs(x = "Term 1", y = "Term 2", fill = "Posterior correlation") +
	scale_fill_distiller(limits = c(-1, 1)) +
	coord_equal() +
	theme(
		legend.position = "top",
		legend.text = element_text(size = 8),
		axis.title = element_blank(),
		axis.text.y = element_text(hjust = 1, size = 7),
		axis.text.x = element_text(angle = 90, hjust = 1, size = 7)
	)

@@

m <- emmeans(model_fits$fit_bcn, ~trial_type:lp, by = "age_group", epred = TRUE) %>% 
	as_tibble() %>% 
	clean_names() %>% 
	mutate_at(vars(emmean:upper_hpd), function(x) ifelse(x < 0, 0, x)) %>% 
	mutate_at(vars(emmean:upper_hpd), function(x) ifelse(x > 1, 1, x))

ggplot(m, aes(age_group, emmean, ymin = lower_hpd, ymax = upper_hpd, colour = trial_type)) +
	facet_wrap(~lp, ncol = 1) +
	# geom_line(aes(group = trial_type), size = 1, position = position_dodge(width = 0.25)) +
	geom_errorbar(size = 1, width = 0, position = position_dodge(width = 0.25)) +
	geom_point(size = 4, position = position_dodge(width = 0.25)) +
	labs(x = "Age (months))", y = "P(target)", colour = "Trial type") +
	scale_y_continuous(limits = c(0, 1), labels = percent)  +
	theme(
		legend.title = element_blank(),
		legend.position = "top"
	)

@@

```{r gaze_time, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
sample_size <- participants %>% 
	filter(valid_participant) %>% 
	count(lp, age_group)

n <- expand_grid(
	lp = c("Monolingual", "Bilingual"),
	age_group = paste(c(21, 25, 30), "months"),
	trial_type = unique(gaze$trial_type),
	time_bin_center = seq(min(gaze$time_bin_center), max(gaze$time_bin_center), 0.5)
)

m <- epred_draws(object = model_fits$fit_bcn, newdata = n, ndraws = 50, re_formula = NA) %>% 
	mutate(.value = logit_to_prob(.epred))

ggplot(m, aes(time_bin_center, .value)) +
	facet_grid(fct_rev(lp)~age_group) +
	geom_hline(yintercept = 0.5, colour = "black") +
	geom_line(aes(group = interaction(age_group, trial_type, lp, .draw),
				  colour = trial_type),
			  alpha = 0.15, size = 0.65) +
	stat_summary(aes(group = trial_type, colour = trial_type),
				 fun = mean, geom = "line", size = 1.25) +
	stat_summary(data = gaze, aes(y = logit_to_prob(logit_adjusted), group = trial_type,
								  colour = trial_type),
				 fun.data = mean_se, geom = "errorbar", size = 0.5, na.rm = TRUE,
				 position = position_dodge(width = 0.3)) +
	geom_text(data = sample_size, aes(x = min(gaze$time_bin_center), y = 0.95, label = paste0("N = ", n)),
			  colour = "white", hjust = 0, size = 5) +
	labs(x = "Time bin (100 ms)", y = "Expected posterior mean PTLT", colour = "Trial type",
		 fill = "Trial type", shape = "Trial type", linetype = "Trial type") +
	scale_color_brewer(palette = "Set1") +
	scale_x_continuous(breaks = min(gaze$time_bin_center):max(gaze$time_bin_center), labels = unique(gaze$time_bin)) +
	scale_y_continuous(limits = c(0, 1), labels = percent) + 
	theme(
		legend.position = "top",
		legend.title = element_blank(),
		axis.text.x = element_text(size = 9),
		panel.grid.major.x = element_blank(),
		panel.grid.minor.x = element_blank()
	)
```

>>>>

# load objects
tar_load_globals()
tar_load(participants)
tar_load(attrition)
tar_load(vocabulary)
tar_load(gaze)
tar_load(fit)

@@

participants <- participants %>% 
	left_join(distinct(attrition, participant, valid_participant)) %>% 
	drop_na(participant) %>% 
	left_join(vocabulary) %>% 
	mutate(
		vocab_size_l1_center = scale(vocab_size_l1)[,1],
		vocab_size_total_center = scale(vocab_size_total)[,1]
	)

@@

```{r fixed_effects_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
draws <- gather_draws(fit, `b_.*`, `sigma`, regex = TRUE)

@@

draws %>% 
	filter(!str_detect(.variable, "time_bin|sigma")) %>% 
	mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4)) %>% 
	ggplot(aes(.value, .variable)) +
	geom_vline(xintercept = 0, colour = "red", size = 1) +
	# stat_halfeye(point_color = "black", interval_color = "black") +
	stat_pointinterval() +
	labs(x = "Value", y = "Variable", fill = "CrI") +
	scale_fill_brewer(direction = -1, na.translate = FALSE) +
	scale_x_continuous(labels = percent) +
	theme(
		axis.title = element_blank(),
		legend.position = "top"
	)

@@

vcov(fit) %>% 
	cov2cor() %>% 
	as.data.frame() %>% 
	rownames_to_column("term1") %>% 
	pivot_longer(-term1, names_to = "term2", values_to = ".value") %>% 
	filter(!grepl("time_bin", term1), !grepl("time_bin", term2)) %>% 
	ggplot(aes(fct_rev(term1), term2, fill = .value)) +
	geom_tile() +
	labs(x = "Term 1", y = "Term 2", fill = "Posterior correlation") +
	scale_fill_distiller(limits = c(-1, 1)) +
	coord_equal() +
	theme(
		legend.position = "top",
		legend.text = element_text(size = 8),
		axis.title = element_blank(),
		axis.text.y = element_text(hjust = 1, size = 7),
		axis.text.x = element_text(angle = 90, hjust = 1, size = 7)
	)

@@

m <- emmeans(fit, ~trial_type:lp, by = "vocab_size_total_center", epred = TRUE) %>% 
	as_tibble() %>% 
	clean_names() %>% 
	mutate_at(vars(emmean:upper_hpd), function(x) ifelse(x < 0, 0, x)) %>% 
	mutate_at(vars(emmean:upper_hpd), function(x) ifelse(x > 1, 1, x))

ggplot(m, aes(vocab_size_total_center, emmean, ymin = lower_hpd, ymax = upper_hpd, colour = trial_type)) +
	facet_wrap(~lp, ncol = 1) +
	# geom_line(aes(group = trial_type), size = 1, position = position_dodge(width = 0.25)) +
	geom_errorbar(size = 1, width = 0, position = position_dodge(width = 0.25)) +
	geom_point(size = 4, position = position_dodge(width = 0.25)) +
	labs(x = "Age (months))", y = "P(target)", colour = "Trial type") +
	scale_y_continuous(limits = c(0, 1), labels = percent)  +
	theme(
		legend.title = element_blank(),
		legend.position = "top"
	)

@@

```{r gaze_time, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
sample_size <- participants %>% 
	filter(valid_participant) %>% 
	count(lp, vocab_size_total_center)

n <- expand_grid(
	lp = c("Monolingual", "Bilingual"),
	vocab_size_total_center = c(-1, 1),
	trial_type = unique(gaze$trial_type),
	time_bin_center = seq(min(gaze$time_bin_center), max(gaze$time_bin_center), 0.1)
)

m <- epred_draws(object = fit, newdata = n, ndraws = 50, re_formula = NA) %>% 
	mutate(
		.value = logit_to_prob(.epred),
		vocab_size_total_center = ifelse(
			vocab_size_total_center==-1, 
			"Low vocabulary size (-1 SD)",
			"High vocabulary size (+1 SD)"
		)
	)

# stat_summary(
# 	aes(group = trial_type, colour = trial_type),
# 	fun = mean, geom = "line", size = 1.25
# ) +
# stat_summary(
# 	data = gaze, aes(
# 		x = as.factor(time_bin_center), y = logit_to_prob(logit_adjusted), 
# 		group = trial_type,
# 		colour = trial_type
# 	),
# 	fun.data = mean_se, geom = "errorbar", size = 0.5, na.rm = TRUE,
# 	position = position_dodge(width = 0.3), inherit.aes = FALSE
# ) 
ggplot(m, aes(time_bin_center, .value)) +
	facet_grid(fct_rev(lp)~vocab_size_total_center) +
	geom_hline(yintercept = 0.5, colour = "black") +
	geom_line(
		aes(group = interaction(vocab_size_total_center, trial_type, lp, .draw),
			colour = trial_type),
		alpha = 0.5, size = 0.65
	) +

	labs(x = "Time bin (100 ms)", y = "Expected posterior mean PTLT", colour = "Trial type",
		 fill = "Trial type", shape = "Trial type", linetype = "Trial type") +
	ggsci::scale_color_futurama() +
	scale_x_continuous(
		breaks = min(gaze$time_bin_center):max(gaze$time_bin_center), 
		labels = 1:20
	) +
	scale_y_continuous(limits = c(0, 1), labels = percent) + 
	theme_minimal() +
	theme(
		legend.position = "top",
		legend.title = element_blank(),
		axis.text.x = element_text(size = 9),
		panel.grid.major.x = element_blank(),
		panel.grid.minor.x = element_blank(),
		panel.background = element_rect(fill = "white", colour = NA),
		plot.background = element_rect(fill = "white", colour = NA)
	)

### 2022-01-13 presentations/2022-01-25_lacre/2022-01-25_lacre-abstract.Rmd
Sample posterior

Previous studies have provided evidence that lexical access is language non-selective in bilinguals: recognising and producing words in one language activates lexical representations of words in the other language [e.g., @costa2000cognate; @thierry2007brain]. It has been suggested that this parallel activation is already present during toddlerhood [e.g., @von2012language; @jardak2019labels], but it is unclear how this parallel activation impacts the developing lexicon. We tested Catalan-Spanish and English-Spanish bilinguals in a word recognition task. Participants were tested exclusively in their language of most exposure. two pictures were presented side-by-side, one of them was named, and participants' visual preference for the named (target) picture was taken as a proxy of word recognition. Each pair of pictures was preceded by the presentation of a silent image (prime). We manipulated two variables regarding to the relationship between the label associated to the prime image and the target word: their phonological overlap in the testing language and the cognate status of the prime label. This resulted in three types of trials: trials in which prime and target labels shared phonological onset and the prime label was a cognate (cognate trials; e.g., *flower*/*flor*-*bird*, for an English-Spanish participant), trials where prime and target labels shared phonological onset but the prime label was a non-cognate (non-cognate trials; e.g., *chair*/*silla*-*bird*), and trials where prime and target labels did not share phonological onset (unrelated trials; *car*/*coche*-*bird*).

In line with previous studies, we expected participants to generate implicit labels for the prime pictures, which should interfere with target word recognition when both words share phonological onset [e.g.; @mani2010infant]. Under the hypothesis that bilingual participants would activate labels for the prime pictures in both languages, we further predicted  that interference over target word recognition would be stronger after cognate primes (labels from both languages interfere with target word recognition) than after non-cognate primes (only the label in the target language interferes with target word recognition). We tested bilinguals (and same-aged monolingual controls) at three age points (21, 25, and 30 months) to investigate how any cross-language priming effect emerged or changed across these ages, capitalising on the potential role of vocabulary size. Preliminary data suggest that bilinguals were almost insensitive to the phonological similarity between the prime and the spoken target word, or to the cognate status of the prime label. Surprisingly, monolinguals' preference for the target picture at 25 was stronger after non-cognate primes, compared to cognate primes or unrelated primes, and also after cognate and non-cognate primes compared to unrelated primes. We present this preliminary data along with exploratory analyses addressing the puzzling pattern of results we observed.

![](img/gca.png)

>>>>

Previous studies suggest that bilinguals access their lexicon in a language non-selective way: recognising and producing words in one language activates word representations in the other [e.g., @costa2000cognate; @thierry2007brain]. This parallel activation of both languages seems to be already present during toddlerhood [e.g., @von2012language], but its impact on the developing lexicon remains unclear. We investigated this issue by testing bilingual toddlers in a primed word recognition task adapted from @mani2010infant. In each trial, participants were first presented with a prime picture in silence. Then, two more pictures were presented side by side, and one of them (target picture) was named in participants' dominant language. We registered participants' looking preference for the target picture as an indicator of word recognition. We designed three types of trials by manipulating the phonological overlap between the prime and the target labels (at word onset), and the cognate status of the prime label. Following @mani2010infant's findings, we expected participants to generate an implicit label for the prime picture, which should interfere with the recognition of the subsequent recognition of the target word if both labels share phonological overlap. Under the hypothesis that bilingual participants would activate labels for the prime pictures in both languages, we further predicted that interference would be stronger after cognate primes (labels from both languages overlap with the target word) than after non-cognate primes (only the label in the language of test overlaps with the target word). We tested bilingual toddlers (and same-aged monolingual controls) at three age points (21, 25, and 30 months) to investigate how parallel activation of both languages emerged or changed across these ages, capitalising on the potential role of vocabulary size. We present preliminary data from this ongoing study, and discuss the potential implications of the current collected data on out hypotheses.
